---
title: "Unit 2"
author: "Richard Palmer"
date: "9/2/2019"
output:
  powerpoint_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo =FALSE}
# get libraries
library(dplyr)
library(ggplot2)

# helpful function
to_inches <- function(height) {
  vals <- unlist(strsplit(x=height,split='-'))
  feet   <- as.numeric(vals[1])
  inches <-  as.numeric(vals[2])
  return (12*feet + inches)
}

# load data
raw_data <- read.csv(file="PlayersBBall.csv")

# convert height to inches
BBall <- raw_data %>% rowwise() %>% mutate(htc=as.character(height),height_inches=to_inches(htc))

```
## Question 1
Show the number of players per position
```{r,echo=TRUE}
BBall %>% filter(position != '')  %>% ggplot() + geom_bar(mapping=aes(x=position)) + ggtitle("Count per position")

```

## Question 2
Compare the distribution of weight of centers with that of the weight of forwards
```{r}
BBall %>% filter(position == 'C' | position == 'F') %>% ggplot(aes(x=weight)) + geom_histogram(bins=20) + ggtitle("Distribution of Weight for Centers and Forwards") + facet_wrap(~position)

```

## Question 3
Compare the distribution of the height of centers with that of the height of forwards
```{r}
BBall %>% filter(position == 'C' | position == 'F') %>% ggplot(aes(x=height_inches)) + geom_histogram(bins=20) + ggtitle("Distribution of Height for Centers and Forwards") + facet_wrap(~position)
```

## Question 4
Investigate distribution of height across all positions
```{r}
BBall %>% filter(position != "") %>% ggplot(aes(x=height_inches)) + geom_histogram(bins=20) + ggtitle("Distribution of Height for all positions") + facet_wrap(~position)
```

## Question 5
Investigate how a players height is related to his weight
```{r}
BBall %>% ggplot(aes(x=weight,y=height_inches)) + geom_point() + ggtitle("Height versus Weight");
```

## Question 6
Is there any difference in height/weight relationship among positions?
```{r}
BBall %>% ggplot(aes(x=weight,y=height_inches,color=position)) + geom_point() + ggtitle("Height versus Weight");
```

## Question 7
Is there any evidence that height has changed over the years?
```{r}
BBall %>% ggplot(aes(x=year_start,y=height_inches)) + geom_point() + ggtitle("Height versus starting year")
```
## Question 8
3D plot of height versus weight, colored by position


## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

