---
title: "Unit6 PreLive"
author: "Richard Palmer"
date: "9/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# get libraries
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stringr)
library(caret)
library(class)
library(XML) #xml_Parse
library(RCurl) #getURL
library(jsonlite)
library(rvest)
```

# Part 1

## Get the data 
```{r}
passengers_url <- "https://public.opendatasoft.com/api/records/1.0/search/?dataset=titanic-passengers&rows=2000&facet=survived&facet=pclass&facet=sex&facet=age&facet=embarked"
passenger_json <- jsonlite::fromJSON(passengers_url,flatten=TRUE)
passenger_df <- passenger_json$records
passenger_valid <- passenger_df %>% select(fields.age,fields.pclass,fields.survived) %>%
  drop_na()
```

## Construct a KNN classifier using age and class
```{r}
#test different values of k.  3 seems pretty good
prediction <- knn(passenger_valid[,1:2],passenger_valid[,1:2],passenger_valid$fields.survived,k=3)
confusionMatrix(prediction,as.factor(passenger_valid$fields.survived))
```

```{r}
myself <- data.frame(fields.age = 56, fields.pclass = c(1,2,3))
prediction <- knn(passenger_valid[,1:2],myself,passenger_valid$fields.survived,k=3,prob=TRUE)
prediction
```

A 56 year old passenger is predicted to survive if in first class, but not if
in second or third class.

## get the list of passengers for whom to predict outcomes and predict
```{r}
titanic_raw <- read.csv('titanic_test.csv')
titanic_test <- titanic_raw %>% rename(fields.age = Age, fields.pclass = Pclass) %>% 
  select(fields.age,fields.pclass) %>% drop_na()
prediction <- knn(passenger_valid[,1:2],titanic_test,passenger_valid$fields.survived)

```


## for a confusion matrix, check predictions on all passengers
```{r}
all_predictions <- knn(passenger_valid[,1:2],passenger_valid[,1:2], passenger_valid$fields.survived,k=3) 
cm <- confusionMatrix(all_predictions,as.factor(passenger_valid$fields.survived))
cm
```

# Part 2

## limit the data
```{r}
of_interest <- iris %>% select(Sepal.Length,Sepal.Width,Species)
```

## parameters
```{r}
set.seed(42)
splitPerc = .7
iterations = 100
numks = 90
```

## run the trials
```{r}
acc_tt = matrix(nrow=iterations,ncol=numks)
for (j in 1:iterations)
{
  trainIndices = sample(1:dim(iris)[1],round(splitPerc * dim(iris)[1]))
  train = of_interest[trainIndices,]
  test  = of_interest[-trainIndices,]
  for (i in 1:numks)
  {
    classifications <- knn(train[,1:2],test[1:2],train[,3],k=i)
    cm = confusionMatrix(classifications,test$Species)
    acc_tt[j,i] = cm$overall[1]
  }
}
```

## analyze the result
```{r}
meanacc <- colMeans(acc_tt)
train_test_df <- as.data.frame(cbind(seq(1,numks,1),colMeans(acc_tt)))
names(train_test_df) <- c("k", "Accuracy")
train_test_df %>% ggplot(aes(x=k,y=Accuracy)) + 
  geom_point() + ggtitle("Accuracy vs k for Train/Test split") + xlab("k") + ylab("Accuracy")
```

## more
```{r}
tt_max_index <- which.max(meanacc)
tt_max_accuracy <- max(meanacc)
```

Using the train/test cross validation, the peak Accuracy of `r tt_max_accuracy`
is found at k=`r tt_max_index`

## run the trials
```{r}
acc_icv = matrix(nrow=iterations,ncol=numks)
for (j in 1:iterations)
{
  for (i in 1:numks)
  {
    classifications <- knn.cv(of_interest[,1:2],of_interest[,3],k=i)
    cm = confusionMatrix(classifications,of_interest$Species)
    acc_icv[j,i] = cm$overall[1]
  }
}
```

## analyze the result
```{r}
meanacc <- colMeans(acc_icv)
plot(seq(1,numks,1),meanacc,type="l")

meanacc <- colMeans(acc_icv)
cv_df <- as.data.frame(cbind(seq(1,numks,1),colMeans(acc_icv)))
names(cv_df) <- c("k", "Accuracy")
cv_df %>% ggplot(aes(x=k,y=Accuracy)) + 
  geom_point() + ggtitle("Accuracy vs k for Leave-One-Out") + xlab("k") + ylab("Accuracy")

```

## more
```{r}
cv_max_index <- which.max(meanacc)
cv_max_accuracy <- max(meanacc)
```

Using leave one out cross validation, the peak Accuracy of `r cv_max_accuracy`
is found at k=`r cv_max_index`
