---
title: "Pre LiveSession 4"
author: "Richard Palmer"
date: "9/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML) #xml_Parse
library(dplyr)
library(tidyr)
library(stringi)
library(rvest) #html_table, html_node
library(ggplot2)
library(RCurl) #getURL
```

## Part 1

### Get xml data from cloudfront.net

```{r}

# use XML package
#restaurants <-   getURL("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%wFrestaurants.xml")
#doc <- xmlParse(restaurants)

# use rvest
#hp <- read_html("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%wFrestaurants.xml")

```

## Part 2

Investigate the quantmod package.

```{r}
library('quantmod')
```

The quantmod package for R is designed to assist the quantitative trader in the development, testing, and deployment of statistically based trading models.
It includes easy to use functions for fetching price data from Google, Yahoo, 
The Federal Reserve, and others.  Although intended as a resource for packages
that would automate trading, it can also be used to simple retrieve financial data.

Fetch historical stock data for Hewlett-Packard.

This function retrieves data into a special 'xts' object.
The argument is the stock symbol, and the name of the resulting
object is the same as that symbol.  It is easy to create a 
data frame from the xts object.
```{r}
getSymbols('HPQ')
hp_data = data.frame(HPQ)
str(hp_data)
summary(hp_data)
```

The quantmod package includes nice plotting functions.
Plot price data using quantmod package
```{r}
chartSeries(HPQ, TA=NULL)
```

I would like to investigate how often HP stock closes on a high
or on a low for the day.  Also, compute the 'closing percentage'
as a number between 0 and 1.  Zero would mean it closes on the low
for the day, 1 means it closes on the high for the day.

```{r}
hp_data <- hp_data %>% 
  mutate(pr = HPQ.High - HPQ.Low) %>%
  mutate(close_on_high = (HPQ.Close == HPQ.High)) %>%
  mutate(close_on_low  = (HPQ.Close == HPQ.Low)) %>%
  mutate(close_perc = (HPQ.Close - HPQ.Low)/pr)

hp_high_closes = sum(hp_data$close_on_high)
hp_low_closes = sum(hp_data$close_on_low)
hp_high_close_rate = 100*hp_high_closes/nrow(hp_data)
hp_low_close_rate = 100*hp_low_closes/nrow(hp_data)
```

Plot a histogram of the 'closing percentage'
```{r}
hp_data %>% ggplot(aes(x=close_perc)) + geom_histogram() + ggtitle('Histogram of closing percentage')
```


It seems that the stock closes on a high only `r hp_high_close_rate` percent of the time, and that it closes on a low `r hp_low_close_rate` percent of the time.



