---
title: "Pre LiveSession 3"
author: "Richard Palmer"
date: "9/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(GGally)
library(dplyr)
library(ggplot2)

```

## Part 1

Load the data and select positions of interest
```{r}
# function to convert height to numeric inches
to_inches <- function(height) {
  vals <- unlist(strsplit(x=height,split="'"))
  feet   <- as.numeric(vals[1])
  inches <-  as.numeric(vals[2])
  return (12*feet + inches)
}
# function to convert string weight to integer
to_pounds <- function(weight) {
  vals <- unlist(strsplit(x=weight,split="lbs"))
  lbs = as.numeric(vals[1])
  return (lbs)
}

# function to convert weight to numeric pounds

fifa_raw <- read.csv("FIFA Players.csv")
fifa_all <- fifa_raw %>% rowwise() %>% mutate(height_inches = to_inches(toString(Height)), weight_pounds = to_pounds(toString(Weight)))

fifa_lm_lf <- fifa_all %>% filter(Position == 'LF' | Position == 'LM')
fifa_lm_lf$Position <- droplevels(fifa_lm_lf$Position)
```

plot Acceleration and Agility versus Position
```{r}
fifa_lm_lf %>% select(Position,Acceleration,Agility) %>% ggpairs(aes(color=Position))
```
### T-test

- Hypotheses

H0: The mean of agility for left midfielders is the same as that for left forwards.
Ha: The mean of agiliety for left midfielders is different from that for left forwards.

- Critical Value



- Test statistic

- p-value

- Accept/Reject

- Conclusion

### Are the assumptions met?

- Normality

- equal variance

- independence

## Part 2



## Takeaways/Questions

### Performance

I wonder how efficient dplyr is.  Am I making lots of copies of data in memory when I have a long string of functions?
